### 2018-03-26 tex/Notation.tex

\begin{tabular}{|l|l|p{10cm}|}
\hline
Type & Notation & Definition \\
\hline

\multirow{10}{3cm}{Queue Specification}

& $\lambda$ & Rate parameter of exponential inter-arrival distribution $f_\delta = \text{Exp}(\lambda)$ for $M/M/K$ queue. \\
& $\mu$ & Rate parameter of exponential service distribution $f_{\mathbf{s}} = \text{Exp}(\mu)$ for $M/M/K$ queue.  \\
& $\rho := \frac{\lambda}{K \mu}$ & Traffic intensity, defined only for $M/M/K$ queues. \\ 
& $\theta_{I}$ & Parameters of arrival and service joint distribution for QDC, $(\mathbf{a}, \mathbf{s}) \sim f_{\mathbf{a}, \mathbf{s}}(\cdot | \theta_I )$. \\

& $K$ & Number of servers. \\
& $C$ & Capacity of system.  \\
& $n$ & Total number of customers. \\
& $R$ & Service discipline of queue. \\
& FCFS & First come first serve (Type of service discipline) \\
& $L$ & Number of knot locations for server changes \\

\hline

\multirow{8}{3cm}{Input/Output} 
& $\mathbf{a} = (a_1, \cdots, a_n )$ & Arrival process, where $a_i$ is the time at which the $i$th customer arrives at the queue. \\

& $\mathbf{\delta} = (\delta_1, \cdots, \delta_n )$ & Inter-arrival process, where $\delta_i$ is $a_i - a_{i-1} \quad \forall i \in 1:n$ and $\delta_1 = a_1$. \\

& $\mathbf{s} = (s_1, \cdots, s_n )$ & Service process, where $s_i$ is the service time of the $i$th customer. \\

& $\mathbf{d} = (d_1, \cdots, d_n )$ & Departure process, where $d_i$ is the time at which the $i$th customer leaves the queue after being served. \\

& $\mathbf{p} = (p_1, \cdots, p_n )$ & Server process, where $p_i$ is the server who served the $i$th customer. \\

& $\mathbf{b} = (b_1, \cdots, b_K )$ & This vector represents the time at which each server $1:K$ will next be free. We consider this vector to be the state of the system. \\

& $\mathbf{x} = (x_1, \cdots, x_L )$ & Change times for number of open servers. \\

& $\mathbf{y} = (y_1, \cdots, y_{L+1} )$ & Number of open servers in each epoch. \\

\hline

\multirow{4}{3cm}{Performance Measures}

& $N(t)$ & Number of customers in system at time $t$. \\
& $\bar{B}$ & Time average number of busy servers, referred to as ``Resource utilization". \\
& $\bar{w}$ & Average waiting time per customer. \\

\hline

\end{tabular}